HO unification from object language to meta language

### 2024-03-28 paper/main.tex
minor

\noindent
The \elpiIn{name} predicate holds only on nominal constants (i.e. bound
variables).\footnote{Elpi provides it as a builtin, but one could implement it by
systematically loading the hypothetical rule \elpiIn{name x} every time
a nominal constant is postulated via \elpiInFN{pi x\ }}
The choice of using n-ary application, rather than binary, is to make it
easy to access the application's head. The price we pay is that substituting
an application in the head of an application should be amended by
``flattening'' \elpiIn{fapp} nodes, that is the job of \elpiIn{napp}.
\footnote{Note that \elpiIn{napp} is an artefact of formalization of \Fo{}
we do in this presentation and, as we explain later,
no equivalent of \elpiIn{napp} is needed in \Ho.}
Finally note
Note that the cut operator is inessential, it could be
removed at the cost of a verbose test on the head of \elpiIn{L}
in the last rule (\elpiIn{L} head
can be \elpiIn{fcon}, \elpiIn{flam} or a name).

@@

The signature of the \elpiIn{comp} predicate below allows for the generation of
links (suspended unification problems) that play no role in this section
but play a major role in Sections \ref{sec:eta} and \ref{sec:beta}.
With respect to \ref{sec:problem-statement} the signature also allows
for updates to the substitution.  The code below only allocates space
for the variables, i.e. sets their memory address to \elpiIn{none},
a details not worth mentioning in the previous discussion.

@@

\noindent
This preliminary version of \elpiIn{comp} recognizes \Fo{} variables
applied to a (possibly empty) duplicate free list of names (i.e.
\elpiIn{pattern-fragment} detects variables in \llambda).
Note tha compiling \elpiIn{Ag} cannot create new mappings nor links, since \elpiIn{Ag}
is made of bound variables and the hypothetical rule loaded by \elpiIn{comp-lam}
(see below) grants this property.

@@

\noindent
Note that we could remove the second rule, whose purpose is to make
links more readable by pruning unneeded abstractions (unused context entries).

@@

Since compilation moves problematic terms out of the compiled terms,
\Ue{} can only perform a partial occur check. For example the
unification problem $X \Ue f~Y$ cannot generate a cyclic substitution alone,
but should be disallowed if a $\linkStore$ contains a link like
$\linketaM{}{Y}{\lambda z.X_z}$: We don't know yet if $Y$ will feature
a lambda in head position, but we surely know it contains $X$, hence
$f~Y$ and that triggers occur check.
The procedure \elpiIn{occur-check-links} is in charge of ensuring that
each link does not represent a (suspended) unification problem doomed
to fail because of occur check. This check is needed in order to
guarantee \cref{prop:fidelity} (simulation fidelity).

>>>>

\noindent
The \elpiIn{name} predicate holds only on nominal constants (i.e. bound
variables).\footnote{Elpi provides it as a builtin, but one could implement it by
systematically loading the hypothetical rule \elpiIn{name x} every time
a nominal constant is postulated via \elpiInFN{pi x\ }}
The choice of using n-ary application, rather than binary, is to make it
easy to access the application's head. The price we pay is that substituting
an application in the head of an application should be amended by
``flattening'' \elpiIn{fapp} nodes, that is the job of \elpiIn{napp}.
\footnote{Note that \elpiIn{napp} is an artefact of formalization of \Fo{}
we do in this presentation and, as we explain later,
no equivalent of \elpiIn{napp} is needed in \Ho.}
Finally note that the cut operator is inessential, it could be
removed at the cost of a verbose test on the head of \elpiIn{L}
in the second rule about \elpiIn{fapp}: \elpiIn{L}'s head
can be \elpiIn{fcon}, \elpiIn{flam} or a name.

@@

The signature of the \elpiIn{comp} predicate below allows for the generation of
links (suspended unification problems) that play no role in this section
but play a major role in \cref{sec:eta} and \cref{sec:beta}.
With respect to \cref{sec:problem-statement} the signature also allows
for updates to the substitution.  The code below uses that possibility
in order to allocate space for the variables, i.e. sets their memory
address to \elpiIn{none} (a details not worth mentioning in the
previous discussion).

@@

\noindent
This preliminary version of \elpiIn{comp} recognizes \Fo{} variables
applied to a (possibly empty) duplicate free list of names.
Note that compiling \elpiIn{Ag} cannot create new mappings nor links, since \elpiIn{Ag}
is made of bound variables and the hypothetical rule loaded by \elpiIn{comp-lam}
(see below) grants this property.

@@

\noindent
Note that we could remove the first rule, whose solve purpose is to make
links more readable by pruning unused context entries.

@@

Since compilation moves problematic terms out of the sigh of \Ue{},
that procedure can only perform a partial occur check. For example the
unification problem $X \Ue f~Y$ cannot generate a cyclic substitution alone,
but should be disallowed if a $\linkStore$ contains a link like
$\linketaM{}{Y}{\lambda z.X_z}$: We don't know yet if $Y$ will feature
a lambda in head position, but we surely know it contains $X$, hence
$f~Y$ and that fails the occur check.
The procedure \elpiIn{occur-check-links} is in charge of ensuring that
each link does not represent a (suspended) unification problem doomed
to fail because of occur check. This check is needed in order to
guarantee \cref{prop:fidelity} (simulation fidelity).
